---
title: "Get Started with yusufHAIGermany"
output:
  rmarkdown::html_vignette:
    toc: true
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Get Started with yusufHAIGermany}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 5.2, fig.align = "center"
)
library(dplyr)
library(tidyr)
library(plotly)
library(lubridate)
```

```{r load-data, message=FALSE}
#load the datasets from the installed package during vignette build
if (!requireNamespace("yusufHAIGermany", quietly = TRUE)) {
  stop("Package not available in vignette build process.")
}

data("sim_monthly", package = "yusufHAIGermany")
data("sim_weekly",  package = "yusufHAIGermany")
data("sim_daily",   package = "yusufHAIGermany")

#quick sanity check so the build fails early if something is missing
stopifnot(
  exists("sim_monthly"), exists("sim_weekly"), exists("sim_daily")
)
```

## Overview

This vignette gives a short tour of the package datasets and shows how to make quick, interactive checks.

The package provides three simulation tables for HAIs in Germany (2011 – 2012):

- sim_monthly: columns date, year, month, hai, cases_month, deaths_month, dalys_month

- sim_weekly: columns date, year, iso_week, hai, cases_week, deaths_week, dalys_week

- sim_daily: columns date, year, yday, wday, hai, cases_day, deaths_day, dalys_day

## Data at a glance

```{r}
dplyr::glimpse(sim_monthly)
dplyr::glimpse(sim_weekly)
dplyr::glimpse(sim_daily)
```

## Monthly totals by infection type

```{r}
p_month <- sim_monthly |>
  mutate(month_lab = format(as.Date(date), "%b %Y")) |>
  group_by(date, hai) |>
  summarise(cases = sum(cases_month), .groups = "drop") |>
  arrange(date) |>
  plot_ly(x = ~date, y = ~cases, 
          color = ~hai, type = "scatter", 
          mode = "lines",
          hovertemplate = paste("%{x|%b %Y}",
                                "%{fullData.name}: %{y:,.0f} cases")
          ) |>
  layout(
    yaxis = list(title = "Cases (monthly)"),
    xaxis = list(title = "Date")
)

p_month
```

## Top infection type by month

```{r}
top_tbl <- sim_monthly |>
  group_by(date, hai) |>
  summarise(cases = sum(cases_month), .groups = "drop") |>
  group_by(date) |>
  slice_max(order_by = cases, n = 1, with_ties = FALSE) |>
  ungroup() |>
  arrange(date)

top_tbl
```

## Annual totals

```{r}
annual <- sim_monthly |>
  mutate(yr = year(as.Date(date))) |>
  group_by(yr) |>
  summarise(
    cases = sum(cases_month),
    deaths = sum(deaths_month),
    dalys = sum(dalys_month),
    .groups = "drop"
    )

annual
```

